---
title: "05 LIME"
date: "2020-11-05"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Load model
```{r}
# LIME FEATURE EXPLANATION ----

# 1. Setup ----

# Load Libraries 

library(h2o)
library(recipes)
library(readxl)
library(tidyverse)
library(tidyquant)
library(lime)
library(rsample)

product_data <- read_csv("./01_ml_fund_source/Business\ Decisions\ with\ Machine\ Learning/product_backorders.csv")
product_data2 <- product_data %>% 
  mutate(
      product_backorder = went_on_backorder %>% str_to_lower() %>% str_detect("yes") %>% as.numeric()
  ) %>% 
  select(-c(went_on_backorder))
glimpse(product_data)

split_obj<- initial_split(product_data2, prop = 0.75)
train_tbl<- training(split_obj)
test_tbl<- testing(split_obj)


# ML Preprocessing Recipe 
recipe_obj <- recipe(product_backorder ~., data = train_tbl) %>% 
    step_zv(all_predictors()) %>% 
    step_dummy(all_nominal(),-all_outcomes()) %>%
    prep()


split_obj<- initial_split(product_data2, prop = 0.75)
train_tbl<- training(split_obj)
test_tbl<- testing(split_obj)

```

```{r}
# 2. Models ----

h2o.init()

automl_leader <- h2o.loadModel("StackedEnsemble_AllModels_AutoML_20210603_202714/StackedEnsemble_AllModels_AutoML_20210603_210724")
automl_leader
```


# Plot
```{r}
# 3.1 Making Predictions ----

predictions_tbl <- automl_leader %>% 
    h2o.predict(newdata = as.h2o(test_tbl)) %>%
    as.tibble() %>%
    bind_cols(
        test_tbl %>%
            select(everything())
    )

predictions_tbl
```

```{r}
summary(train_tbl)
```

Take the explanation data and use the first case to create a plot similar to the output of plot_features().

## Original plot_features()
```{r}
# explanation %>% 
#   as.tibble()
#   
# case_1 <- explanation %>%
#     filter(case == 1)
# 
# case_1 %>%
#     plot_features()
# You will need at least the layers geom_col() and coord_flip().

explainer <- train_tbl %>%
    select(-product_backorder) %>%
    lime(
        model           = automl_leader,
        bin_continuous  = TRUE,
        n_bins          = 4,
        quantile_bins   = TRUE
    )

explainer


explanation <- test_tbl %>%
    slice(1) %>%
    select(-product_backorder) %>%
    lime::explain(
    
        # Pass our explainer object
        explainer = explainer,
        # Because it is a binary classification model: 1
        n_labels   = 1,
        # number of features to be returned
        n_features = 8,
        # number of localized linear models
        n_permutations = 5000,
        # Let's start with 1
        kernel_width   = 1
    )

explanation
```

```{r}

g <- plot_features(explanation = explanation, ncol = 1, cases = 1)
g
```


## Bonus Objectives:
Get your custom plot_features() function to scale to multiple cases
Use theme arguments to modify the look of the plot

```{r}
explanation_multi <- test_tbl %>%
    slice(1:20) %>%
    select(-product_backorder) %>%
    lime::explain(
        explainer = explainer,
        n_labels   = 1,
        n_features = 8,
        n_permutations = 5000,
        kernel_width   = 0.5
    )

explanation_multi %>%
    as.tibble()

plot_explanations(explanation_multi)

```

# Part 2: Recreate plot_explanations():

Take the full explanation data and recreate the second plot.
You will need at least the layers geom_tile() and facet_wrap().

## Customized plot_features()
```{r}
theme_lime <- function(...) {
  theme_minimal() +
    theme(
      strip.text = element_text(face = 'bold', size = 9),
      plot.margin = margin(15, 15, 15, 15),
      legend.background = element_blank(),
      legend.key = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_blank(),
      legend.position = 'bottom',
      panel.spacing.y = unit(15, 'pt'),
      strip.text.x = element_text(margin = margin(t = 2, b = 2), hjust = 0),
      axis.title.y = element_text(margin = margin(r = 10)),
      axis.title.x = element_text(margin = margin(t = 10)),
      panel.background = element_rect(fill   = "transparent"),
      panel.border     = element_rect(color = "black", fill = NA, size = 0.5),
      panel.grid.major = element_line(color = "grey", size = 0.333),
      ...
    )
}

plot_explanations_customized <- function(explanation, ...) {

  num_cases <- unique(suppressWarnings(as.numeric(explanation$case)))
  if (!anyNA(num_cases)) {
    explanation$case <- factor(explanation$case, levels = as.character(sort(num_cases)))
  }
  explanation$feature_desc <- factor(
    explanation$feature_desc,
    levels = rev(unique(explanation$feature_desc[order(explanation$feature, explanation$feature_value)]))
  )
  
p <- ggplot(explanation, aes_(~case, ~feature_desc),show.legend=TRUE) +
    geom_tile(aes_(fill = ~feature_weight)) +
    scale_x_discrete('Case', expand = c(0, 0)) +
    scale_y_discrete('Feature', expand = c(0, 0)) +
    scale_fill_gradient2('Feature\nweight', low = 'firebrick', mid = '#f7f7f7', high = 'steelblue') +
    theme_lime() +
    theme(panel.border = element_rect(fill = NA, colour = 'grey60', size = 1),
          panel.grid = element_blank(),
          legend.position = 'right',
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  if (is.null(explanation$label)) {
    p
  } else {
    p + facet_wrap(~label, ...)
  }
}

```

```{r}
plot_explanations_customized(explanation = explanation, ncol = 1, cases = 1)
```

```{r}
plot_explanations_customized(explanation_multi)
```